\paper. \textit{Optimizing the Bruck Algorithm for Non-uniform All-to-all Communication.}
Ke Fan, \textbf{Thomas Gilray}, Valerio Pascucci, Xuan Huang, Kristopher Micinski, and Sidharth Kumar.
International ACM Symposium on High-Performance Parallel and Distributed Computing.
\\(HPDC---19\% acceptance) Jun 2022. 
\\ \vspace{-0.1cm}\\
\paper. \textit{Seamless Deductive Inference Via Macros.}
Arash Sahebolamri, \textbf{Thomas Gilray}, and Kristopher Micinski.
International Conference on Compiler Construction.
\\(CC) Apr 2022. 
\\ \vspace{-0.1cm}\\
\paper. \textit{Load-balancing Parallel I/O of Compressed Hierarchical Layouts.}
Ke Fan, Duong Hoang, Steve Petruzza, \textbf{Thomas Gilray}, Valerio Pascucci, and Sidharth Kumar.
International Conference on High Performance Computing, Data, and Analytics.
\\(HiPC) Dec 2021. 
\\ \vspace{-0.1cm}\\
\paper. \textit{Compiling Data-parallel Datalog.}
\textbf{Thomas Gilray}, Sidharth Kumar, and Kristopher Micinski.
International Conference on Compiler Construction.
\\(CC) Mar 2021. 
\\ \vspace{-0.1cm}\\
\paper. \textit{Load-balancing Parallel Relational Algebra.}
Sidharth Kumar and \textbf{Thomas Gilray}.
ISC High Performance.
\\(ISC---31\% acceptance) Jun 2020. 
\\\textbf{Won ISC Hans Meuer Best Paper award.} \\ \vspace{-0.1cm}\\
\paper. \textit{Abstracting Faceted Execution.}
Kristopher Micinski, David Darais, and \textbf{Thomas Gilray}.
IEEE Computer Security Foundations Symposium.
\\(CSF) Jun 2020. 
\\ \vspace{-0.1cm}\\
\paper. \textit{Distributed Relational Algebra at Scale.}
Sidharth Kumar and \textbf{Thomas Gilray}.
International Conference on High Performance Computing, Data, and Analytics.
\\(HiPC---23\% acceptance) Dec 2019. 
\\\textbf{Won HiPC Best Paper award.} \\ \vspace{-0.1cm}\\
\paper. \textit{Size-Change Termination as a Contract.}
Phúc C. Nguyễn, \textbf{Thomas Gilray}, Sam Tobin-Hochstadt, and David Van Horn.
Programming Language Design and Implementation.
\\(PLDI---27\% acceptance) Jun 2019. 
\\\textbf{Invited to the Journal of Functional Programming.} \\ \vspace{-0.1cm}\\
\paper. \textit{Soft Contract Verification for Higher-order Stateful Programs.}
Phúc C. Nguyễn, \textbf{Thomas Gilray}, Sam Tobin-Hochstadt, and David Van Horn.
Symposium on Principles of Programming Languages.
\\(PoPL---23\% acceptance) Jan 2018. 
\\ \vspace{-0.1cm}\\
\paper. \textit{User Comfort with Android Background Resource Accesses in Different Contexts.}
Daniel Votipka, Seth M. Rabin, Kristopher Micinski, \textbf{Thomas Gilray}, Michelle L. Mazurek, and Jeffrey S. Foster.
Symposium on Usable Privacy and Security.
\\(SOUPS---21\% acceptance) 2018. 
\\ \vspace{-0.1cm}\\
\paper. \textit{Allocation Characterizes Polyvariance.}
\textbf{Thomas Gilray}, Michael D. Adams, and Matthew Might.
International Conference on Functional Programming.
\\(ICFP---31\% acceptance) Sep 2016. 
\\ \vspace{-0.1cm}\\
\paper. \textit{Pushdown Control-Flow Analysis for Free.}
\textbf{Thomas Gilray}, Steven Lyde, Michael D. Adams, Matthew Might, and David Van Horn.
Symposium on Principles of Programming Languages.
\\(PoPL---23\% acceptance) Jan 2016. 
\\ \vspace{-0.1cm}\\
\paper. \textit{Dynamic Sparse-Matrix Allocation on GPUs.}
James King, \textbf{Thomas Gilray}, Robert M. Kirby, and Matthew Might.
ISC High Performance.
\\(ISC) 2016. 
\\\textbf{Won PRACE ISC best paper award.} \\ \vspace{-0.1cm}\\
\paper. \textit{A Survey of Polyvariance in Abstract Interpretations.}
\textbf{Thomas Gilray} and Matthew Might.
Symposium on Trends in Functional Programming.
\\(TFP) May 2013. 
\\\textbf{Won TFP best student paper award.} \\ \vspace{-0.1cm}\\
