\paper. \textit{Datalog with First-class Facts.}
\textbf{Thomas Gilray}, Arash Sahebolamri$\ddagger$, Yihao Sun$\ddagger$, Sowmith Kunapaneni$\dagger$, Sidharth Kumar, and Kristopher Micinski.
International Conference on Very Large Data Bases.
\\(VLDB) Sep 2025.
\\ \vspace{-0.1cm}\\
\paper. \textit{Optimizing Datalog for the GPU.}
Yihao Sun$\ddagger$, Ahmedur Rahman Shovon$\ddagger$, \textbf{Thomas Gilray}, Sidharth Kumar, and Kristopher Micinski.
International Conference on Architectural Support for Programming Languages and Operating Systems.
\\(ASPLOS---12.7\% acceptance) Mar 2025. \url{https://doi.org/10.1145/3669940.3707274}
\\ \vspace{-0.1cm}\\
\paper. \textit{Column-Oriented Datalog on the GPU.}
Yihao Sun$\ddagger$, Sidharth Kumar, \textbf{Thomas Gilray}, and Kristopher Micinski.
AAAI Conference on Artificial Intelligence.
\\(AAAI---23.4\% acceptance) Feb 2025.
\\ \vspace{-0.1cm}\\
\paper. \textit{Optimizing the Bruck Algorithm for Non-uniform All-to-all Communication.}
Ke Fan$\ddagger$, \textbf{Thomas Gilray}, Valerio Pascucci, Xuan Huang, Kristopher Micinski, and Sidharth Kumar.
International ACM Symposium on High-Performance Parallel and Distributed Computing.
\\(HPDC---19\% acceptance) Jun 2022. \url{https://doi.org/10.1145/3502181.35314}
\\ \vspace{-0.1cm}\\
\paper. \textit{Load-balancing Parallel Relational Algebra.}
Sidharth Kumar and \textbf{Thomas Gilray}.
ISC High Performance.
\\(ISC---31\% acceptance) Jun 2020. \url{https://doi.org/10.1007/978-3-030-50743-5_15}
\\\textbf{Won ISC Hans Meuer Best Paper award.} \\ \vspace{-0.1cm}\\
\paper. \textit{Distributed Relational Algebra at Scale.}
Sidharth Kumar and \textbf{Thomas Gilray}.
International Conference on High Performance Computing, Data, and Analytics.
\\(HiPC---23\% acceptance) Dec 2019. \url{https://doi.org/10.1109/HiPC.2019.00014}
\\\textbf{Won HiPC Best Paper award.} \\ \vspace{-0.1cm}\\
\paper. \textit{Size-Change Termination as a Contract.}
Phúc C. Nguyễn$\ddagger$, \textbf{Thomas Gilray}, Sam Tobin-Hochstadt, and David Van Horn.
Programming Language Design and Implementation.
\\(PLDI---27\% acceptance) Jun 2019. \url{https://doi.org/10.1145/3314221.3314643}
\\\textbf{Invited to the Journal of Functional Programming.} \\ \vspace{-0.1cm}\\
\paper. \textit{Soft Contract Verification for Higher-order Stateful Programs.}
Phúc C. Nguyễn$\ddagger$, \textbf{Thomas Gilray}, Sam Tobin-Hochstadt, and David Van Horn.
Symposium on Principles of Programming Languages.
\\(POPL---23\% acceptance) Jan 2018. \url{https://doi.org/10.1145/3158139}
\\ \vspace{-0.1cm}\\
\paper. \textit{Allocation Characterizes Polyvariance.}
\textbf{Thomas Gilray}, Michael D. Adams, and Matthew Might.
International Conference on Functional Programming.
\\(ICFP---31\% acceptance) Sep 2016. \url{https://doi.org/https://doi.org/10.1145/3022670.2951936}
\\ \vspace{-0.1cm}\\
\paper. \textit{Pushdown Control-Flow Analysis for Free.}
\textbf{Thomas Gilray}, Steven Lyde, Michael D. Adams, Matthew Might, and David Van Horn.
Symposium on Principles of Programming Languages.
\\(POPL---23\% acceptance) Jan 2016. \url{https://doi.org/10.1145/2837614.2837631}
\\ \vspace{-0.1cm}\\
\paper. \textit{Dynamic Sparse-Matrix Allocation on GPUs.}
James King, \textbf{Thomas Gilray}, Robert M. Kirby, and Matthew Might.
ISC High Performance.
\\(ISC) 2016. \url{https://doi.org/10.1007/978-3-319-41321-1_4}
\\\textbf{Won PRACE ISC best paper award.} \\ \vspace{-0.1cm}\\
